[Re] A neural model of the saccade generator in the reticular formation

### 2018-02-07 article/senden-schuecker-hahne-diesmann-goebel-2018.tex
updated equations

The SG model described by Gancarz \& Grossberg \autocite{Gancarz1998}
consists of a horizontal and a vertical component each with two
long-lead burst neurons (LLBNs), excitatory burst neurons (EBNs),
inhibitory burst neurons (IBNs), and tonic neurons (TNs). Within each
component, the two directions (left-right, down-up) interact
antagonistically. Additionally, both components share a single omnipause
neuron (OPN) which tonically inhibits each EBN to suppress unwanted
saccades. In what follows we briefly describe the dynamics of neurons
within the horizontal component (equivalent descriptions apply to the
vertical component).

@@

\[
\begin{array}{ll}
\tau\frac{dL_l}{dt} &= -1.3L_l+I_l-2B_l \\\\
\tau\frac{dL_r}{dt} &= -1.3L_r+I_r-2B_r \\
\end{array}
\]

@@

\[
\begin{array}{ll}
\tau\frac{dE_l}{dt} &= -3.5E_l+(2-E_l)(5E_l+1)-(E_l+1)(10L_r+20g(P))) \\\\
\tau\frac{dE_r}{dt} &= -3.5E_r+(2-E_r)(5E_r+1)-(E_r+1)(10L_l+20g(P))) \\
\end{array}
\] with a nonlinear gain function \(g(x)\) given by

\[
g(x) = \frac{x^4}{0.1^4+x^4} \textrm{.}
\]

@@

\[
\begin{array}{ll}
\tau\frac{dB_l}{dt} &= -2.4B_l+3E_l \\\\
\tau\frac{dB_r}{dt} &= -2.4B_r+3E_r \\
\end{array}
\]

@@

\[
\tau\frac{dP}{dt} = -0.2P+(1-P)(1.2+J)-3.5(P+0.4)(g(L_l)+g(L_r)+g(L_u)+g(L_d))
\] where \(J\) represents external electrical stimulation.

@@

\[
\begin{array}{ll}
\tau\frac{dT_l}{dt} &= 0.1(E_l-E_r) \\\\
\tau\frac{dT_r}{dt} &= 0.1(E_r-E_l) \textrm{.} \\
\end{array}
\] Horizontal eye position (\(\theta\)) depends on activity of the right
TN; i.e. \(\theta = 260(T_r-0.5)\).

@@

In all simulations we used the Exponential Euler method for numerical
integration of rate neurons \autocite{Hahne2016} at a time step of
\(0.05\,\mathrm{ms}\) and a time constant of \(50\,\mathrm{ms}\). It
should be noted that rates of all neurons were initialized to zero and
were thus not at resting equilibrium. In order to address this, the
model was allowed to evolve for \(100\,\mathrm{ms}\), to relax towards
equilibrium before applying any input. Furthermore, we always simulated
the full model; i.e.~both its horizontal and vertical components even if
input was applied only to one of the two.

@@

The seventh simulation shows that saccade velocity and duration can be
traded while keeping amplitude constant. To produce a high-velocity
saccade, the SC was stimulated at a high frequency. Similarly, to
produce a low-velocity saccade, the SC was stimulated at a low
frequency. Figure \ref{fig:fig_7} shows the results of our simulation.
In line with the original publication, the saccade amplitude reflected
by TN activity was identical after high- and low-frequency saccades thus
confirming the reported effect. However, we observe differences in the
specific details of our implementation with respect to the original,
starting with the shape of the input curves. While the rise of the input
curves in the original publication and our implementation coincide well,
the decay in our curves started later and proceeded with a shorter time
constant (\(\approx25\,\mathrm{ms}\), i.e.~half the simulation time
constant). This resulted in more total input to the model in our
implementation and hence lead to the generation of a second saccade not
observed in the original publication. When repeating our simulation
using an ad-hoc solution to create matching input curves (by halving the
time constant after external stimulation), we observed results identical
to those reported in the original publication. We proceeded to
investigate the origin of the discrepancy with regard to the shapes of
the input curves. First, we tested whether discrepancies were specific
to our NEST implementation by solving the expressions describing the
input analytically. Specifically, according to equations A13-A15 of
\autocite{Gancarz1998} the input \(I\) is described by \[
\begin{array}{lll}
    \tau \frac{dA}{dt} &=-A+F(t)\\
    f(x) &=
    \left\{
    \begin{array}{lll}
        0 \quad &\textrm{if} \quad x<0 \\
        x \quad &\textrm{if} \quad 0<x<1 \\
        1 \quad &\textrm{if} \quad x \geq 1 \\
    \end{array}
    \right. \\
    I(t) &=Wf(A)
\end{array}
\]\\
with the stimulus \[
F(t)=
\left\{
\begin{array}{lll}
    0 \quad &\textrm{if} \quad t<{t}_{on} \\
    F \quad &\textrm{if} \quad {t}_{on}<t<{t}_{off} \textrm{.} \\
    0 \quad &\textrm{if} \quad t \geq {t}_{off} \\
\end{array}
\right.
\]\\
Given this, the analytic expression for \(I(t)\) is \begin{equation}
I(t)=
\left\{
\begin{array}{lll}
    0 \quad &\textrm{if} \quad t<{t}_{on} \\
    W \cdot f(F \cdot (1-{e}^{-(t-{t}_{on})/ \tau})) \quad &\textrm{if} \quad {t}_{on}<t<{t}_{off} \textrm{.}\\
    W \cdot f(F \cdot (1-{e}^{-({t}_{off}-{t}_{on})/ \tau}) \cdot {e}^{-(t-{t}_{off})/ \tau}) \quad &\textrm{if} \quad t \geq {t}_{off} \\
\end{array}
\right.
\label{eq:1}\end{equation} Evaluating the analytic expression given
parameter values reported in the original manuscript exactly reproduced
our numerical results and thus produced curves equally deviatng from
those shown in the original publication. Next, we contacted one of the
original authors to rule out the possibility that we used erroneous
parameter settings but no mistake was pointed out to us. The origin of
the discrepancy thus remains unclear. However, it is likely it stems
from the description of the input rather than from problems with the SG
model itself.

\begin{figure}
\centering
\includegraphics[width=0.75000\textwidth,height=0.75000\textwidth]{figures/fig7.eps}
\caption{\textbf{Trading saccade velocity for duration.} Duration and
velocity of a saccade can be traded while keeping amplitude constant. To
produce a high velicity saccade an input of \(\mathrm{F=3}\) was applied
to the SC for \(68\,\mathrm{ms}\) (\(82\,\mathrm{ms}\) in the original
publication). To produce a low velicity saccade an input of
\(\mathrm{F=1.3}\) for \(117\,\mathrm{ms}\). Given the shape of the
input curve produced by our implementation of the original equations
(A), a second saccade can be observed for both high and low velocity
saccades. If the input curves reported in the original manuscript were
recreated in an ad-hoc fashion (B), the responses of all model neurons
matched those in \autocite{Gancarz1998}\label{fig:fig_7}}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.60000\textwidth,height=0.46200\textwidth]{figures/fig8.eps}
\caption{\textbf{Smooth staircase eye movements.} Activity profiles of
SG neurons accompanying smooth eyes movement as a result of strong
(\(\mathrm{I=3}\)) sustained (\(300\,\mathrm{ms}\))
input.\label{fig:fig_8}}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.46200\textwidth,height=0.70000\textwidth]{figures/fig9.eps}
\caption{\textbf{Interrupted saccade resulting from OPN stimulation.}
OPN stimulation interrupted the saccade, which remained accurate
nonetheless. An input of (\(\mathrm{I=0.7}\)) was applied to the LLBN
for \(100\,\mathrm{ms}\). At \(45\,\mathrm{ms}\) after onset of the
input, the OPN was stimulated (\(\mathrm{J=1.8}\)) for
\(5\,\mathrm{ms}\). The dashed line shows TN activity for an
uninterrupted saccade.\label{fig:fig_9}}
\end{figure}

\clearpage

>>>>

The SG model described by Gancarz \& Grossberg \autocite{Gancarz1998}
consists of a horizontal and a vertical component each with two
long-lead burst neurons (LLBNs), excitatory burst neurons (EBNs),
inhibitory burst neurons (IBNs), and tonic neurons (TNs). Within each
component, the two directions (left-right, down-up) interact
antagonistically. Additionally, both components share a single omnipause
neuron (OPN) which tonically inhibits each EBN to suppress unwanted
saccades. While this constitutes the core SG model, the superior
collicuus (SC) is relevant for some simulations. In what follows we
briefly describe the dynamics of neurons within the horizontal component
(equivalent descriptions apply to the vertical component).

@@

\begin{equation}
\begin{array}{ll}
\tau\frac{dL_l}{dt} &= -1.3L_l+I_l-2B_l \\\\
\tau\frac{dL_r}{dt} &= -1.3L_r+I_r-2B_r \\
\end{array}
\label{eq:llbn}\end{equation}

@@

\begin{equation}
\begin{array}{ll}
\tau\frac{dE_l}{dt} &= -3.5E_l+(2-E_l)(5E_l+1)-(E_l+1)(10L_r+20g(P))) \\\\
\tau\frac{dE_r}{dt} &= -3.5E_r+(2-E_r)(5E_r+1)-(E_r+1)(10L_l+20g(P))) \\
\end{array}
\label{eq:ebn}\end{equation} with a nonlinear gain function \(g(\cdot)\)
given by

\begin{equation}
g(x) = \frac{x^4}{0.1^4+x^4} \textrm{.}
\label{eq:gain}\end{equation}

@@

\begin{equation}
\begin{array}{ll}
\tau\frac{dB_l}{dt} &= -2.4B_l+3E_l \\\\
\tau\frac{dB_r}{dt} &= -2.4B_r+3E_r \\
\end{array}
\label{eq:ibn}\end{equation}

@@

\begin{equation}
\tau\frac{dP}{dt} = -0.2P+(1-P)(1.2+J)-3.5(P+0.4)(g(L_l)+g(L_r)+g(L_u)+g(L_d))
\label{eq:opn}\end{equation} where \(J\) represents external electrical
stimulation and \(g(\cdot)\) again given by equation \ref{eq:gain}.

@@

\begin{equation}
\begin{array}{ll}
\tau\frac{dT_l}{dt} &= 0.1(E_l-E_r) \\\\
\tau\frac{dT_r}{dt} &= 0.1(E_r-E_l) \textrm{.} \\
\end{array}
\label{eq:tn}\end{equation} Horizontal eye position (\(\theta\)) depends
on activity of the right TN; i.e. \(\theta = 260(T_r-0.5)\).

\subsubsection{Superior colliculus}\label{superior-colliculus}

The superior colliculus (\(A\)) receives electrical stimulation F

\begin{equation}
\tau \frac{dA}{dt} =-A+F(t) \textrm{.}
\label{eq:sc}\end{equation}

\subsubsection{SC input to the saccade
generator}\label{sc-input-to-the-saccade-generator}

For a range of simulations, external input to the saccade generator is
obtained by applying a nonlinearity \(f(\cdot)\) given by
\begin{equation}
f(x) = 
\left\{
 \begin{array}{lll}
    0 \quad &\textrm{if} \quad x<0 \\
    x \quad &\textrm{if} \quad 0<x<1 \\
    1 \quad &\textrm{if} \quad x \geq 1 \\
  \end{array}
\right.\
\label{eq:pw}\end{equation} to the activity \(A\) of the SC and scaling
the result by a weight \(W\) such that

\begin{equation}
I = Wf(A)\textrm{.}
\label{eq:stim}\end{equation}

@@

In all simulations we used the Exponential Euler method for numerical
integration of rate neurons \autocite{Hahne2017} at a time step of
\(0.05\,\mathrm{ms}\) and a time constant of \(50\,\mathrm{ms}\). It
should be noted that rates of all neurons were initialized to zero and
were thus not at resting equilibrium. In order to address this, the
model was allowed to evolve for \(100\,\mathrm{ms}\), to relax towards
equilibrium before applying any input. Furthermore, we always simulated
the full model; i.e.~both its horizontal and vertical components even if
input was applied only to one of the two.

@@

\begin{figure}
\centering
\includegraphics[width=0.90000\textwidth,height=0.90000\textwidth]{figures/fig7.eps}
\caption{\textbf{Trading saccade velocity for duration.} Duration and
velocity of a saccade can be traded while keeping amplitude constant. To
produce a high velicity saccade an input of \(\mathrm{F=3}\) was applied
to the SC for \(68\,\mathrm{ms}\) (\(82\,\mathrm{ms}\) in the original
publication). To produce a low velicity saccade an input of
\(\mathrm{F=1.3}\) for \(117\,\mathrm{ms}\). Given the shape of the
input curve produced by our implementation of the original equations
(A), a second saccade can be observed for both high and low velocity
saccades. If the input curves reported in the original manuscript were
recreated in an ad-hoc fashion (B), the responses of all model neurons
matched those in \autocite{Gancarz1998}\label{fig:fig_7}}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.70000\textwidth,height=0.56200\textwidth]{figures/fig8.eps}
\caption{\textbf{Smooth staircase eye movements.} Activity profiles of
SG neurons accompanying smooth eyes movement as a result of strong
(\(\mathrm{I=3}\)) sustained (\(300\,\mathrm{ms}\))
input.\label{fig:fig_8}}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.46200\textwidth,height=0.60000\textwidth]{figures/fig9.eps}
\caption{\textbf{Interrupted saccade resulting from OPN stimulation.}
OPN stimulation interrupted the saccade, which remained accurate
nonetheless. An input of (\(\mathrm{I=0.7}\)) was applied to the LLBN
for \(100\,\mathrm{ms}\). At \(45\,\mathrm{ms}\) after onset of the
input, the OPN was stimulated (\(\mathrm{J=1.8}\)) for
\(5\,\mathrm{ms}\). The dashed line shows TN activity for an
uninterrupted saccade.\label{fig:fig_9}}
\end{figure}

The seventh simulation shows that saccade velocity and duration can be
traded while keeping amplitude constant. To produce a high-velocity
saccade, the SC was stimulated at a high frequency. Similarly, to
produce a low-velocity saccade, the SC was stimulated at a low
frequency. Figure \ref{fig:fig_7} shows the results of our simulation.
In line with the original publication, the saccade amplitude reflected
by TN activity was identical after high- and low-frequency saccades thus
confirming the reported effect. However, we observe differences in the
specific details of our implementation with respect to the original,
starting with the shape of the input curves. While the rise of the input
curves in the original publication and our implementation coincide well,
the decay in our curves started later and proceeded with a shorter time
constant (\(\approx25\,\mathrm{ms}\), i.e.~half the simulation time
constant). This resulted in more total input to the model in our
implementation and hence lead to the generation of a second saccade not
observed in the original publication. When repeating our simulation
using an ad-hoc solution to create matching input curves (by halving the
time constant after external stimulation), we observed results identical
to those reported in the original publication. We proceeded to
investigate the origin of the discrepancy with regard to the shapes of
the input curves. First, we tested whether discrepancies were specific
to our NEST implementation by solving the expressions describing the
input analytically. Specifically, the analytic expression of the input
\(I(t)\) described by equations \ref{eq:sc} - \ref{eq:stim} is
\begin{equation}
I(t)=
\left\{
\begin{array}{lll}
    0 \quad &\textrm{if} \quad t<{t}_{on} \\
    W \cdot f(F \cdot (1-{e}^{-(t-{t}_{on})/ \tau})) \quad &\textrm{if} \quad {t}_{on}<t<{t}_{off} \textrm{.}\\
    W \cdot f(F \cdot (1-{e}^{-({t}_{off}-{t}_{on})/ \tau}) \cdot {e}^{-(t-{t}_{off})/ \tau}) \quad &\textrm{if} \quad t \geq {t}_{off} \\
\end{array}
\right.
\label{eq:analytic}\end{equation} Evaluating the analytic expression
given parameter values reported in the original manuscript exactly
reproduced our numerical results and thus produced curves equally
deviating from those shown in the original publication. Next, we
contacted one of the original authors to rule out the possibility that
we used erroneous parameter settings but no mistake was pointed out to
us. The origin of the discrepancy thus remains unclear. However, it is
likely it stems from the description of the input rather than from
problems with the SG model itself.

@@
